Deterministic Scheduling of Periodic Messages for Cloud RAN

### 2017-07-21 infocom/infocom.tex
debut aprem

\newcommand{\todo}[1]{{\color{red} TODO: {#1}}}

%opening
\title{Deterministic Scheduling of Periodic Messages by an SDN Approach}

@@

\begin{abstract}
A recent trend in mobile networks is to centralize the processing units which were attached to 
antennas in distant data-centers. The main challenge is to guarantee that the latency of the periodic messages 
sent from the antennas to their processing units and back is minimal. We show that traditional statistical multiplexing
is bad at ensuring a low latency. Hence, we propose in this article to use a deterministic scheme to send periodic messages without collisions in the network and thus without buffering and additional latency. 
We give several algorithms to compute such schemes for a common topology where one link is shared by all antennas.
We show that there is always a solution when the routes are small or the load is small. When the parameters are unconstrained,
and some buffering is allowed in the processing units, we propose an algorithm (PMLS) adapted from a classical scheduling method.
The experimental results shows that even under full load, most of the time PMLS finds a deterministic sending scheme with no latency.
\end{abstract}

@@

Our main contributions are the following.
 In Sec.~\ref{sec:def} we propose a model of the network and the periodic sending of messages along its routes
 and we introduce the star routed network we study in this article. 
 We formalize the problem of finding a periodic assignment for sending messages without buffering (PAZL) or
 with buffering in the processing unit only (PALL) and shows that they are $\NP$-hard even to approximate.
In Sec.~\ref{sec:PAZL}, we propose two algorithms solving PAZL when the routes between RRHs and BBUs are small
or when the load is light. Using a compact representation of an assignment, we show that PAZL is fixed parameter tractable when parametrized by the number of antenna on star routed networks.
Finally in Sec.~\ref{sec:PALL}, we introduce a family of algorithms to solve PALL and our experimentation 
allows to select one which works extremely well even in very loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional statistical multiplexing with regard to latency.

@@

The network is modeled as a directed graph $G=(V,A)$. Each arc  $(u,v)$ in $A$ is labeled by an integer weight $\omega(u,v)$ which represents the time taken by a signal to go from $u$ to $v$ using this arc. A {\bf route} $r$ in $G$ is a path, that is, a sequence of adjacent vertices $u_0, \ldots , u_{k}$, with $(u_i,u_{i+1}) \in A$.  The {\bf latency} of a vertex $u_i$ in a path $r$ is defined by $\lambda(u_i,r)= \sum\limits_{0 \leq j <i} \omega(u_j, u_{j+1})$.
% We also define $\lambda(u_0,r)=0$. 
The length of the route $r$ is defined by $\lambda (r)= \lambda (u_k,r)$.
We denote by $\cal R$ a set of routes, the pair $(G,\cal R)$ is called a {\bf routed network} and represents our telecommunication network. The first vertex of a route models an antenna (RRH) and the last one a data-center (BBU) which processes the messages sent by the antenna.

@@

@@

@@

      \noindent {\bf Periodic Assignment for Low Latency (PALL)}

      In Sec.~\ref{sec:PALL} we will study algorithms which solve the search version of PALL (computing an assignment), also denoted by PALL for simplicity. \todo{phrase utile finalement ?} 
      It turns out that the problem PALL is hard to solve for \emph{general} routed networks. The long version of this article contains complete proofs of the $\NP$-hardness or hardness of approximation of PALL and of several of its variants.

 \begin{theorem}
Problem PALL is $\NP$-complete.
\end{theorem}

\noindent \emph{Sketch of proof:}
 We first show that finding an assignment for the forward routes only with no collisions 
 can be reduced to PALL by a symmetry argument.
 We fix $\tau$ the size of messages to $1$ and we reduce the problem of $k$-coloring to the 
 problem of finding an assignment without collisions for the forward routes.
 Let $H$ be a graph instance of the $k$-coloring problem, the reduction is as follows. 
 For each vertex $v$ in $H$, there is a route $r_v$ in ${\cal R}$. Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an arc in $H$; this arc is the only one shared by these two routes. All arcs are of weight $0$. 
 Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $(k,1)$-periodic assignment in $G$, 
 by converting an offset of a route into a color of a vertex and reciprocally. \qed

       We consider a simpler version of PALL, that we call {\bf Periodic Assignment for Zero Latency} or PAZL: we ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$}. We introduce PAZL because it is simpler to study and we are able to prove theoretical results and find better algorithms than for PALL. Moreover, as we show in our experimentations of Sec.\ref{sec:exp_PAZL}, this problem can often be solved positively (albeit less often than the general problem). Finally, a solution to PAZL is simpler to implement in real telecommunication networks, since we do not need to implement any buffering at all.

       We were not able to prove that PAZL is $\NP$-hard and we leave this question open for further work. 
       It seems very likely to be hard, since a variant of PAZL where the central link is unidirectional (collisions can happen between forward and backward routes) can be shown to be $\NP$-complete. It is done by a reduction from the subset sum problem as for a similar problem of scheduling pair of tasks~\cite{orman1997complexity}.

@@

\section{Solving PAZL on Star Routed Networks} \label{sec:PAZL}

  In this section, we deal with the problem PAZL on star routed networks. Since we want an assignment with waiting times zero, $MT({\cal M})$ is equal to twice the length of the longest route, thus $T_{max}$ is not relevant anymore. Moreover choosing the offset $m_r$ of a forward route $r$ also sets the offset of the backward route $\rho(r)$ to $m_{\rho(r)} = m_{r} + \lambda(r) \mod P$.  Consider an assignment $m$ of a star routed network and let $m'_r= m_{r} - t(c_s,r) \mod P$.
  Then $m'$ is an assignment of the same star routed network where, for all $i$, the weights of the arcs from $s_i$ to $c_s$ are set to $0$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

\begin{theorem}\label{th:FPT}
$PALL \in \FPT$ when parametrized by the number of routes.
\end{theorem}
\begin{proof}
Let $(G, {\cal R})$ be a star routed network and let $\mathcal{M}$ be a $(P,\tau)$-periodic assignment of $(G, {\cal R})$. First remark that for a given assignment and a route $r_0$ with offset $m$, by removing $m$ to all offsets, we can always assume that its offset is zero. Therefore we need only to consider all \emph{compact assignments} with an \emph{offset $0$} for the route $r_0$. 
We now evaluate the number of compact assignments and prove that it only depends
on $n$ the number of routes which proves the theorem. 
We give a way to build a compact assignment $\mathcal{M}$ by determining its offsets one after the other. We fix an arbitrary total order on ${\cal R}$.
First a route $r_0$ is chosen arbitrarily and its offset set to $0$. 
Then at each step, if the offsets of $S \subseteq  {\cal R}$ have been chosen,
we select the smallest route $r$ in $S$ for the order. 
Then we select a route in $r' \in {\cal R} \setminus S$ and set its offset such that 
if we remove $1$ then $r'$ collides with $r$. Note that if $r$ is a forward route (resp. a backward route) then $r'$ is also a forward route (resp. a backward route). We can also decide to 
not use route $r$. At a given step of the algorithm, if $|S| = 2i$, we have $n-i$ choices 
of routes to select. The value of the offset of the selected route is entirely determined by the values of the offsets of routes in $S$. Therefore there are at most $n!$ different compact assignments with offset $r_0$ fixed to $0$.

The algorithm to solve PALL builds every possible compact assignment as described here, and
tests at each step whether there is a collision which can be done in time linear in the size of 
$(G, {\cal R})$. Therefore $PALL \in \FPT$.
\end{proof}

@@

      In our experiment, we generate $1000$ random instances of PAZL for $1$ to $12$ routes. 
      We measure the average minimal period, for each algorithm. Remind that the exhaustive search always finds the best possible period for a given instance. The lower and upper bound $n\tau$ and $3n\tau$ are also represented.

      \todo{présenter ça comme une average load ?}

      \begin{center}
	 \hspace{-0.5cm}\includegraphics[width=0.53\textwidth]{periode_petite.pdf}
      \end{center}
            \vspace{-1cm}
      \caption{Minimum period averaged over $1000$ random instances}
      \end{figure}

      First, we remark that the period found by the exhaustive search is only very slightly above the lower bound of $n\tau$, which means that, in this regime, it is very well justified to look for a solution without waiting time even for a highly loaded network.

@@

      When the load is larger than $50\%$ (period less than $40000$), the exhaustive search finds  more solutions than the greedy algorithms which justifies its use. However, for load larger than $80\%$ (period less than $25000$) there are many instances for which there are no solutions to PAZL.
      It means that with long routes and high load, looking for an assignment without waiting time is far too restrictive. That is why we present algorithms for the general PALL problem in our next section. We will test them on $8$ long routes and a load between $100\%$ and $80\%$ for which, as shown here, there are often no assignment without waiting times.

@@

         \section{Solving PALL on Star Routed Networks}\label{sec:PALL}

    In this section, we still work on star routed networks, but we consider again the more general PALL problem. It means that the messages can wait in the target vertices (BBUs) in order to find solutions when there are none without waiting time. We allow the process times $PT$ of the routes to be greater than twice the weights of the routes, but it should less than $T_{max}$.

@@

	Often in real networks, the length of the routes are not arbitrary and we may exploit that to solve PALL easily. For instance if all the weights on the arcs $(c_t,t_i)$ are the same, we can replace them all by $0$ and substract this weight to $T_{max}$. It corresponds to a situation where all the BBUs are in the same datacenter.
	The assignment in that case is trivial, just send all messages so that they follow each other without gaps in the central arc $(c_s,c_t)$. Since the arcs $(c_t,t_i)$ are of weight $0$, all messages will go through $(c_t,c_s)$ on their way back in
	the same order and thus will not collide.

@@

    We can decompose any algorithm solving the PALL problem in two parts: first set the offsets of the forward routes and then, knowing this information set the offset of the backward routes or equivalently the waiting times.  
    The offsets of the forward routes are chosen so that messages have no collision on the central arc and such that there are no free slots in between the end of a message on the central arc and the beginning of the next one. 
    It is done to minimize the period needed to send the messages of the forward route. For this section and the next, we assume that the offsets of the forward routes are fixed by some order.

@@

%     \begin{algorithm}[H]
%     \caption{ Minimized Scheduling Periodic (MSP)}
%     \begin{algorithmic}
%     \REQUIRE A routed network $(G,{\cal R})$,a period $P$, packet size $\tau$, $ T_{max}$, the offsets $m_i$
%     \ENSURE $(P-\tau)-$periodic assignment of $(G,{\cal R})$, if it exists
%   
%     \FORALL{route $r_{t_i}$}
%     \STATE  $w_i \leftarrow 0$
%     \STATE period-end $\leftarrow m_{s_i} + \lambda(r_{s_i}) + t(c_t,r_{t_i}) + P$
%     \FORALL{route $r_{t_j}$}
%     \STATE deadline-route$ \leftarrow m_{s_j} + T_{max}-t(c_s,r_{s_j})$
%     \STATE $deadline \leftarrow$ min(deadline-route,period-end)
%     \ENDFOR
%     
%     \STATE Call (MS)
% 
%     
%     \ENDFOR
% 
%     \STATE return the best $(P,\tau)$-periodic assignment, or FAILURE
% 
%     \end{algorithmic}
%     \end{algorithm}

    \subsection{Experimental evaluation}
    \label{sec:resultsPALL}

    In this section, we first try to understand what is the best choice of order for the first stage of the algorithm.
    We fix the number of routes to $8$ to make comparisons with the results of Section~\ref{sec:exp_PAZL} easier. 
    We draw uniformly the weights of the arcs between $0$ and $20000$ slots.
    For a given routed network, we define the {\em margin} as the difference between $T_{max}$ and twice the longest route. 
    Hence, for a routed network choosing the margin or $T_{max}$ is the same, but since we will consider many networks 
    whith different longest route's size, it will be more meaningful to set the margin rather than $T_{max}$.
    In fact the margin represents the latency imposed by the communication process without taking into account the physical length of the network which cannot be changed. In our experiments the margin range from  $0$ to $3000$ slots.
   We look at two different regimes: we take a period equal to $1,05n\tau$ to represent a very loaded network
   and $1,5n\tau$ for a mildly loaded network. Choosing a larger period is not relevant since then we know how to solve the problem even without waiting times as shown in Section~\ref{sec:exp_PAZL}. We represent the success rate with regards to the margin, for the five kind of orders using the Greedy algorithm for the second stage, computed over $10000$ random star routed networks. Note that for the random order, we compute $1000$ random orders and count it as a success as soon as there is a solution for one order.

@@

  \centering
     \hspace{-1.5cm} \includegraphics[width=0.4\textwidth]{departs_gp_21000.pdf}
\vspace{-0.75cm}
      \caption{Success rate of different orders, $95\%$ load.}
          \end{figure}
          \vspace{-1cm}
\begin{figure}[H] 
          \includegraphics[width=0.5\textwidth]{departs_gp_25000.pdf}
          \vspace{-1.5cm}
      \caption{Success rate of different orders, $80\%$ load.}
           \label{fig:success30000}
     \end{figure}

@@

     The random order is the best by far. When the load is not too high ($P = 30000$) there is always a solution with margin $0$.
     Note that, when disallowing waiting times, there were instances without solutions for these parameters (see Section~\ref{sec:exp_PAZL}), which justifies the interest of studying the PALL problem. We now want to compare the performances of the three different algorithms used in the second stage. Since GD already showed excellent results on mild loads, it is more interesting to focus on the behavior of the algorithms on high load. Moreover, we will use $1000$ random orders for the first stage as it gives the best results. The following experiment is realized on routed networks with $8$ routes, weights of the arcs drawn between $0$ and $20000$ and with a period of $21000$.  We represent the success rate with regards to the margin, for the three algorithms computed over $10000$ random star routed networks.

    \begin{figure} [H] 
       \begin{center}
      \includegraphics[width=0.5\textwidth]{retour_21000.pdf}
      \end{center}
            \vspace{-1cm}
      \caption{Success rate of GD, MLS and PMLS for $P = 21000$}
     \label{fig:success21000}
     \end{figure}

     The MLS algorithm performs poorly, worst than GD and PMLS, which shows that taking into account the periodicity is fundamental.
     The GD algorithms is close to $100\%$ success rate for margins larger than $1500$ while the PMLS algorithm finds a solution in more than $95\%$ of the experiments, even with a margin $0$. Therefore, for the worst possible constraints on load and margin, there are a few instances for which we do not find a solution. However, with a margin of $1000$, which corresponds to about $0.05$ms of additional delay with the chosen parameters, we always find a solution.

    Remark that the random order corresponds to the best of $1000$ orders drawn uniformly which means that a computation can be up to $1000$ time slower than for a fixed order. The choice of the number of random orders drawn yields a trade-off between the computation time and the success rate. Up to now, we chose $1000$ random orders arbitrarily. We investigate the success rate of our algorithm with regards to the number of random orders drawn, a period of $21000$ and a margin $0$. The following tables presents the average success rate for each number of sending orders, on $1000$ instances, for GD and PMLS.

         \begin{figure}[H] 
       \begin{center}
   \begin{tabularx}{0.5\textwidth}{|c|X|X|X|X|X|X|}
    \hline
    \# orders& $1$ & $10$ & $100$& $1 000$& $10 000$&$100 000$\\
    \hline
    GD & $0.6$ &$7.1$&$33.4$&$76.4$&$90.0$&$90.0$\\
    \hline
  PMLS & $41,1$ &$82,5$&$94,5$&$96,2$&$97,2$&$97,2$\\
    \hline
      \end{tabularx}
      \end{center}
   \caption{Impact of the number of random sending orders}
     \end{figure}

First remark that we can improve our previous results by taking $10000$ random orders, which 
yields the best results for the two algorithms. The number of different orders is $7!= 5040$ since we have $8$ routes and the solutions are invariant up to a circular permutation of the order. Therefore instead of doing $10000$ random draws we could test every possible order in less time. However this method would not be practical with regard to the computation time for $n = 15$. On the other hand, remark that the success rate of PMLS is already high for $100$ random orders, which means it will work even better than the other methods when $n$ is larger and that the number of random orders drawn is critical.

     Now that we have found the best amongst the algorithms solving PALL, we need to compare its performances against the actual way to manage the messages in a network. The classical way to manage messages in the network is statistical multiplexing, where there is a FIFO buffer in each node of the network to deal with the collisions. The time at which the messages are sent in the network is not computed as in our approach, thus we fix the offsets of each route to some random value.
     Even if this policy seems to work in practice when the network is not too loaded, it does not give any guarantee on the latency. Remark that the process is not periodic, therefore we must measure the process time of each route over several periods if we want to compute its maximum. We choose to simulate it for $1000$ periods and we have observed that the process time usually stabilizes after $10$ periods. The margin is defined as the maximum process time, computed as explained, minus twice the size of the longest route.

>>>>

\newcommand{\todo}[1]{{\color{red} TODO: {#1}}}
\newcommand\pazl{\textsc{pazl}\xspace}
\newcommand\pall{\textsc{pall}\xspace}
%opening
\title{Deterministic Scheduling of Periodic Messages for Cloud RAN}

@@

\begin{abstract}
A recent trend in mobile networks is to centralize the processing units which were attached to 
antennas in distant data-centers. The main challenge is to guarantee the maximum latency of the periodic messages 
sent from the antennas to their processing units and back. We show that traditional statistical multiplexing
is bad at ensuring a low latency. Hence, we propose in this article to use a deterministic scheme to send periodic messages without collisions in the network and thus without buffering and additional latency. 
We give several algorithms to compute such schemes for a common topology where one link is shared by all antennas.
We show that there is always a solution when the routes are small or the load is small. When the parameters are unconstrained,
and some buffering is allowed in the processing units, we propose an algorithm (PMLS) adapted from a classical scheduling method.
The experimental results show that even under full load, most of the time PMLS finds a deterministic sending scheme with no latency.
\end{abstract}

@@

Our main contributions are the following.
 In Sec.~\ref{sec:def} we propose a model of the network and the periodic sending of messages along its routes
 and we introduce the star routed network we study in this article. 
 We formalize the problem of finding a periodic assignment for sending messages without buffering (\pazl) or
 with buffering in the processing unit only (\pall) and shows that they are $\NP$-hard even to approximate.
In Sec.~\ref{sec:PAZL}, we propose two algorithms solving \pazl when the routes between RRHs and BBUs are small
or when the load is light. Using a compact representation of an assignment, we show that \pazl is fixed parameter tractable when parametrized by the number of antenna on star routed networks.
Finally in Sec.~\ref{sec:PALL}, we introduce a family of algorithms to solve \pall and our experimentation 
allows to select one which works extremely well even in very loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional statistical multiplexing with regard to latency.

@@

The network is modeled as a directed graph $G=(V,A)$. Each arc  $(u,v)$ in $A$ is labeled by an integer weight $\omega(u,v)$ which represents the time taken by a message to go from $u$ to $v$ using this arc. A {\bf route} $r$ in $G$ is a path, that is, a sequence of adjacent vertices $u_0, \ldots , u_{k}$, with $(u_i,u_{i+1}) \in A$.  The {\bf latency} of a vertex $u_i$ in a path $r$ is defined by $\lambda(u_i,r)= \sum\limits_{0 \leq j <i} \omega(u_j, u_{j+1})$.
% We also define $\lambda(u_0,r)=0$. 
The length of the route $r$ is defined by $\lambda (r)= \lambda (u_k,r)$.
We denote by $\cal R$ a set of routes, the pair $(G,\cal R)$ is called a {\bf routed network} and represents our telecommunication network. The first vertex of a route models an antenna (RRH) and the last one a data-center (BBU) which processes the messages sent by the antenna.

@@

      \noindent {\bf Periodic Assignment for Low Latency (\pall)}

      In Sec.~\ref{sec:PALL} we will study algorithms which solve the search version of \pall (computing an assignment), also denoted by \pall for simplicity. \todo{phrase utile finalement ?} 
      It turns out that the problem \pall is hard to solve for \emph{general} routed networks. The long version of this article contains complete proofs of the $\NP$-hardness or hardness of approximation of \pall and of several of its variants.

 \begin{theorem}
Problem \pall is $\NP$-complete.
\end{theorem}

\noindent \emph{Sketch of proof:}
 We first show that finding an assignment for the forward routes only with no collisions 
 can be reduced to \pall by a symmetry argument.
 We fix $\tau$ the size of messages to $1$ and we reduce the problem of $k$-coloring to the 
 problem of finding an assignment without collisions for the forward routes.
 Let $H$ be a graph instance of the $k$-coloring problem, the reduction is as follows. 
 For each vertex $v$ in $H$, there is a route $r_v$ in ${\cal R}$. Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an arc in $H$; this arc is the only one shared by these two routes. All arcs are of weight $0$. 
 Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $(k,1)$-periodic assignment in $G$, 
 by converting an offset of a route into a color of a vertex and reciprocally. \qed

       We consider a simpler version of \pall, that we call {\bf Periodic Assignment for Zero Latency} or \pazl: we ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$}. We introduce \pazl because it is simpler to study and we are able to prove theoretical results and find better algorithms than for \pall. Moreover, as we show in our experimentations of Sec.\ref{sec:exp_PAZL}, this problem can often be solved positively (albeit less often than the general problem). Finally, a solution to \pazl is simpler to implement in real telecommunication networks, since we do not need to implement any buffering at all.

       We were not able to prove that \pazl is $\NP$-hard and we leave this question open for further work. 
       It seems very likely to be hard, since a variant of \pazl where the central link is unidirectional (collisions can happen between forward and backward routes) can be shown to be $\NP$-complete. It is done by a reduction from the subset sum problem as for a similar problem of scheduling pair of tasks~\cite{orman1997complexity}.

@@

  Even if this topology seems really basics, it's already complex to find some solutions for \pazl and \pall on it. Also, a star routed network corresponds to a Point-to-Multipoint fronthaul (PtMP), intended for C-RAN \cite{tayq2017real}.

@@

@@

\section{Solving \pazl on Star Routed Networks} \label{sec:PAZL}

  In this section, we deal with the problem \pazl on star routed networks. Since we want an assignment with waiting times zero, $MT({\cal M})$ is equal to twice the length of the longest route, thus $T_{max}$ is not relevant anymore. Moreover choosing the offset $m_r$ of a forward route $r$ also sets the offset of the backward route $\rho(r)$ to $m_{\rho(r)} = m_{r} + \lambda(r) \mod P$.  Consider an assignment $m$ of a star routed network and let $m'_r= m_{r} - t(c_s,r) \mod P$.
  Then $m'$ is an assignment of the same star routed network where, for all $i$, the weights of the arcs from $s_i$ to $c_s$ are set to $0$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

\begin{theorem}\label{th:FPT}
$\pall \in \FPT$ when parametrized by the number of routes.
\end{theorem}
\begin{proof}
Let $(G, {\cal R})$ be a star routed network and let $\mathcal{M}$ be a $(P,\tau)$-periodic assignment of $(G, {\cal R})$. First remark that for a given assignment and a route $r_0$ with offset $m$, by removing $m$ to all offsets, we can always assume that its offset is zero. Therefore we need only to consider all \emph{compact assignments} with an \emph{offset $0$} for the route $r_0$. 
We now evaluate the number of compact assignments and prove that it only depends
on $n$ the number of routes which proves the theorem. 
We give a way to build a compact assignment $\mathcal{M}$ by determining its offsets one after the other. We fix an arbitrary total order on ${\cal R}$.
First a route $r_0$ is chosen arbitrarily and its offset set to $0$. 
Then at each step, if the offsets of $S \subseteq  {\cal R}$ have been chosen,
we select the smallest route $r$ in $S$ for the order. 
Then we select a route in $r' \in {\cal R} \setminus S$ and set its offset such that 
if we remove $1$ then $r'$ collides with $r$. Note that if $r$ is a forward route (resp. a backward route) then $r'$ is also a forward route (resp. a backward route). We can also decide to 
not use route $r$. At a given step of the algorithm, if $|S| = 2i$, we have $n-i$ choices 
of routes to select. The value of the offset of the selected route is entirely determined by the values of the offsets of routes in $S$. Therefore there are at most $n!$ different compact assignments with offset $r_0$ fixed to $0$.

The algorithm to solve \pall builds every possible compact assignment as described here, and
tests at each step whether there is a collision which can be done in time linear in the size of 
$(G, {\cal R})$. Therefore $\pall \in \FPT$.
\end{proof}

@@

      In our experiment, we generate $1000$ random instances of \pazl for $1$ to $12$ routes. 
      We measure the average minimal period, for each algorithm. Remind that the exhaustive search always finds the best possible period for a given instance. The lower and upper bound $n\tau$ and $3n\tau$ are also represented.
      \todo{présenter ça comme une average load ?}
      \begin{figure}[h]
      \begin{center}
	 \hspace{-0.5cm}\includegraphics[width=0.5\textwidth]{periode_petite.pdf}
      \end{center}
            \vspace{-0.75cm}
      \caption{Minimum period averaged over $1000$ random instances}
      \end{figure}
      First, we remark that the period found by the exhaustive search is only very slightly above the lower bound of $n\tau$, which means that, in this regime, it is very well justified to look for a solution without waiting time even for a highly loaded network.

@@

      When the load is larger than $50\%$ (period less than $40000$), the exhaustive search finds  more solutions than the greedy algorithms which justifies its use. However, for load larger than $80\%$ (period less than $25000$) there are many instances for which there are no solutions to \pazl.
      It means that with long routes and high load, looking for an assignment without waiting time is far too restrictive. That is why we present algorithms for the general \pall problem in our next section. We will test them on $8$ long routes and a load between $100\%$ and $80\%$ for which, as shown here, there are often no assignment without waiting times.

@@

         \section{Solving \pall on Star Routed Networks}\label{sec:PALL}

    In this section, we still work on star routed networks, but we consider again the more general \pall problem. It means that the messages can wait in the target vertices (BBUs) in order to find solutions when there are none without waiting time. We allow the process times $PT$ of the routes to be greater than twice the weights of the routes, but it should less than $T_{max}$.

@@

	Often in real networks, the length of the routes are not arbitrary and we may exploit that to solve \pall easily. For instance if all the weights on the arcs $(c_t,t_i)$ are the same, we can replace them all by $0$ and substract this weight to $T_{max}$. It corresponds to a situation where all the BBUs are in the same datacenter.
	The assignment in that case is trivial, just send all messages so that they follow each other without gaps in the central arc $(c_s,c_t)$. Since the arcs $(c_t,t_i)$ are of weight $0$, all messages will go through $(c_t,c_s)$ on their way back in
	the same order and thus will not collide.

@@

    We can decompose any algorithm solving the \pall problem in two parts: first set the offsets of the forward routes and then, knowing this information set the offset of the backward routes or equivalently the waiting times.  
    The offsets of the forward routes are chosen so that messages have no collision on the central arc and such that there are no free slots in between the end of a message on the central arc and the beginning of the next one. 
    It is done to minimize the period needed to send the messages of the forward route. For this section and the next, we assume that the offsets of the forward routes are fixed by some order.

@@

%     \begin{algorithm}[H]
%     \caption{ Minimized Scheduling Periodic (MSP)}
%     \begin{algorithmic}
%     \REQUIRE A routed network $(G,{\cal R})$,a period $P$, packet size $\tau$, $ T_{max}$, the offsets $m_i$
%     \ENSURE $(P-\tau)-$periodic assignment of $(G,{\cal R})$, if it exists
%   
%     \FORALL{route $r_{t_i}$}
%     \STATE  $w_i \leftarrow 0$
%     \STATE period-end $\leftarrow m_{s_i} + \lambda(r_{s_i}) + t(c_t,r_{t_i}) + P$
%     \FORALL{route $r_{t_j}$}
%     \STATE deadline-route$ \leftarrow m_{s_j} + T_{max}-t(c_s,r_{s_j})$
%     \STATE $deadline \leftarrow$ min(deadline-route,period-end)
%     \ENDFOR
%     
%     \STATE Call (MS)
% 
%     
%     \ENDFOR
% 
%     \STATE return the best $(P,\tau)$-periodic assignment, or FAILURE
% 
%     \end{algorithmic}
%     \end{algorithm}
    \subsection{Experimental evaluation}
    \label{sec:resultsPALL}
    In this section, we first try to understand what is the best choice of order for the first stage of the algorithm.
    We fix the number of routes to $8$ to make comparisons with the results of Section~\ref{sec:exp_PAZL} easier. 
    We draw uniformly the weights of the arcs between $0$ and $20000$ slots.
    For a given routed network, we define the {\em margin} as the difference between $T_{max}$ and twice the longest route. 
    Hence, for a routed network choosing the margin or $T_{max}$ is the same, but since we will consider many networks 
    whith different longest route's size, it will be more meaningful to set the margin rather than $T_{max}$.
    In fact the margin represents the latency imposed by the communication process without taking into account the physical length of the network which cannot be changed. In our experiments the margin range from  $0$ to $3000$ slots.
   We look at two different regimes: we take a period equal to $1,05n\tau$ to represent a very loaded network
   and $1,5n\tau$ for a mildly loaded network. Choosing a larger period is not relevant since then we know how to solve the problem even without waiting times as shown in Section~\ref{sec:exp_PAZL}. We represent the success rate with regards to the margin, for the five kind of orders using the Greedy algorithm for the second stage, computed over $10000$ random star routed networks. Note that for the random order, we compute $1000$ random orders and count it as a success as soon as there is a solution for one order.

@@

  \centering
     \hspace{-1.5cm} \includegraphics[width=0.4\textwidth]{departs_gp_21000.pdf}
\vspace{-0.75cm}
      \caption{Success rate of different orders, $95\%$ load.}
          \end{figure}
          \vspace{-1cm}
\begin{figure}[H] 
          \includegraphics[width=0.4\textwidth]{departs_gp_25000.pdf}
          \vspace{-0.5cm}
      \caption{Success rate of different orders, $80\%$ load.}
           \label{fig:success30000}
     \end{figure}

@@

     The random order is the best by far. When the load is not too high ($P = 30000$) there is always a solution with margin $0$.
     Note that, when disallowing waiting times, there were instances without solutions for these parameters (see Section~\ref{sec:exp_PAZL}), which justifies the interest of studying the \pall problem. We now want to compare the performances of the three different algorithms used in the second stage. Since GD already showed excellent results on mild loads, it is more interesting to focus on the behavior of the algorithms on high load. Moreover, we will use $1000$ random orders for the first stage as it gives the best results. The following experiment is realized on routed networks with $8$ routes, weights of the arcs drawn between $0$ and $20000$ and with a period of $21000$.  We represent the success rate with regards to the margin, for the three algorithms computed over $10000$ random star routed networks. 
    \begin{figure} [H] 
      \vspace{-0.5cm}
       \begin{center}
      \includegraphics[width=0.5\textwidth]{retour_21000.pdf}
      \end{center}
            \vspace{-1cm}
      \caption{Success rate of GD, MLS and PMLS for $P = 21000$}
     \label{fig:success21000}
       \vspace{-0.2cm}
     \end{figure}
     The MLS algorithm performs poorly, worst than GD and PMLS, which shows that taking into account the periodicity is fundamental.
     The GD algorithms is close to $100\%$ success rate for margins larger than $1500$ while the PMLS algorithm finds a solution in more than $95\%$ of the experiments, even with a margin $0$. Therefore, for the worst possible constraints on load and margin, there are a few instances for which we do not find a solution. However, with a margin of $1000$, which corresponds to about $0.05$ms of additional delay with the chosen parameters, we always find a solution.

    Remark that the random order corresponds to the best of $1000$ orders drawn uniformly which means that a computation can be up to $1000$ time slower than for a fixed order. The choice of the number of random orders drawn yields a trade-off between the computation time and the success rate. Up to now, we chose $1000$ random orders arbitrarily. We investigate the success rate of our algorithm with regards to the number of random orders drawn, a period of $21000$ and a margin $0$. The following tables presents the average success rate for each number of sending orders, on $1000$ instances, for GD and PMLS.
         \begin{figure}[H] 
       \begin{center}
   \begin{tabularx}{0.5\textwidth}{|c|X|X|X|X|X|X|}
    \hline
    \# orders& $1$ & $10$ & $100$& $1 000$& $10 000$&$100 000$\\
    \hline
    GD & $0.6$ &$7.1$&$33.4$&$76.4$&$90.0$&$90.0$\\
    \hline
  PMLS & $41,1$ &$82,5$&$94,5$&$96,2$&$97,2$&$97,2$\\
    \hline
      \end{tabularx}
      \end{center}
   \caption{Impact of the number of random sending orders}
     \end{figure}
First remark that we can improve our previous results by taking $10000$ random orders, which 
yields the best results for the two algorithms. The number of different orders is $7!= 5040$ since we have $8$ routes and the solutions are invariant up to a circular permutation of the order. Therefore instead of doing $10000$ random draws we could test every possible order in less time. However this method would not be practical with regard to the computation time for $n = 15$. On the other hand, remark that the success rate of PMLS is already high for $100$ random orders, which means it will work even better than the other methods when $n$ is larger and that the number of random orders drawn is critical.

     Now that we have found the best amongst the algorithms solving \pall, we need to compare its performances against the actual way to manage the messages in a network. The classical way to manage messages in the network is statistical multiplexing, where there is a FIFO buffer in each node of the network to deal with the collisions. The time at which the messages are sent in the network is not computed as in our approach, thus we fix the offsets of each route to some random value.
     Even if this policy seems to work in practice when the network is not too loaded, it does not give any guarantee on the latency. Remark that the process is not periodic, therefore we must measure the process time of each route over several periods if we want to compute its maximum. We choose to simulate it for $1000$ periods and we have observed that the process time usually stabilizes after $10$ periods. The margin is defined as the maximum process time, computed as explained, minus twice the size of the longest route.
